---
title: "hw_2_review"
author: "Theodore Dounias"
date: "February 24, 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(Lahman)
```

**Exercise 3.1**
```{r, echo = FALSE, include = FALSE}
library(mosaic)
```
We create the desired graph
```{r}
ggplot(Galton, aes(father, height)) +
  geom_point(aes(col = sex)) +
  geom_smooth(method = "lm") +
  facet_wrap(~sex)
```

**Exercise 3.2**
```{r}
#Transform RailTrail weekday into more understandable form
RailTrail1 <- RailTrail
bool <- c("no", "yes")
RailTrail1$weekday <- bool[RailTrail$weekday]
#Plot graph
ggplot(RailTrail1, aes(hightemp, volume, col = weekday)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~weekday)
```

**Exercise 4.8**
We create an operational version of Teams, and add two variables using  the mutate function:
```{r}
Teams_op <- Teams %>%
  mutate(BA = H/AB, SLG = (H + X2B + 2*X3B + 3*HR)/AB)
```

**Exercise 4.9**
We first create the appropriatelly wrangled data set and then plot the graph:
```{r}
#Wrangle data
Team_4.9 <- Teams_op %>%
  filter(yearID >=1954)
#Graph
ggplot(Team_4.9, aes(x = yearID, y = SLG, col = lgID)) +
  geom_point(alpha = .3) +
  geom_smooth()
```
It appears that the American League has had typically higher SLG's than the National League. ~~whatever those are~~

**Exercise 4.10**
First for MLB history:
```{r}
Top_teams <- Teams_op %>%
  group_by(teamID) %>%
  summarize(meanSLG = mean(SLG)) %>%
  arrange(desc(meanSLG)) %>%
  slice(1:15)
Top_teams
```
And for recent history:
```{r}
Top_teams_recent <- Teams_op %>%
  filter(yearID >=1969) %>%
  group_by(teamID) %>%
  summarize(meanSLG = mean(SLG)) %>%
  arrange(desc(meanSLG)) %>%
  slice(1:15)
Top_teams_recent
```

**Exercise 4.11**
Assuming that the world series is the most important thing in baseball, and that success of a season depends on rank, the following gives us the wanted results:
```{r}
#Create Dataset
LATeam <- Teams_op %>%
  filter(teamID %in% c("LAA", "CAL", "ANA")) 
#Filter and print for "success" as measured by someone who does not know baseball
LATeam$WSWin <- as.factor(LATeam$WSWin)
LATeam %>%
  select(Rank, yearID, teamID, WSWin) %>%
  arrange(desc(WSWin), Rank) %>%
  slice(1:10)
```


